---
title: "PA1"
author: "kwerkies"
date: "Sunday, April 19, 2015"
output: html_document
---

###Description of Assignment

Using data from a personal activity monitoring device, we seek to understand an individual's activity patterns. Data collected (total of 17568 obs): measured data = steps taken; over 2 months (Oct to Nov 2012); collected at 5-min intervals throughout the day.

####Task 1: Load and Preprocess the data
```{r echo=TRUE}
setwd("C:/Users/weijie/Desktop/Coursera R course/V. Reproducible Research")
fileUrl <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
temp <- tempfile()
download.file(fileUrl,temp)
data <- read.csv(unz(temp,"activity.csv"))
data[,2] <- as.Date(data[,2])

#I construct another dataset, daydata, to represent data (total no. of steps taken) for each day.
n <- 24*60/5
grp <- seq(1,nrow(data),by=n)
daysteps <- sapply(grp, function(x) sum(data[x:(x+n-1),1]))
daydata <- matrix(nrow=length(grp),ncol=2)
colnames(daydata)<-c("total day's steps","date")
for(i in 1:length(grp)){
        daydata[i,1] <- daysteps[i]
        daydata[i,2] <- as.character(data[grp[i],2])
}
```

####Task 2: Calculating statistics on the total number of steps taken per day
```{r, echo=TRUE}
#a. Calculate the total no. of steps taken per day -- Already calculated as daydata dataset.

#b. Make a histogram of the total no. of steps taken per day
hist(as.numeric(daydata),xlab="",main="Histogram of the total number of steps taken per day")

#c. Calculate the mean and median of the total no. of steps taken per day
meandaysteps = round(mean(daysteps,na.rm=TRUE),digits=3)
mediandaysteps=median(daysteps,na.rm=TRUE)
```

The mean and median of the total number of steps taken per day are `r meandaysteps` and `r mediandaysteps` respectively.

####Task 3: Determining the average daily activity pattern
```{r echo=TRUE}
intervaldata <- matrix(nrow=n,ncol=2)
colnames(intervaldata)<-c("interval","average steps at this interval")
for(i in 1:288){
        intervaldata[i,2]=sum(data[grp[]+i-1,1],na.rm=TRUE)
        intervaldata[i,1]=data[i,3]
}
plot(intervaldata,type="l")
hiint = intervaldata[which.max(intervaldata[,2]),1]
```

The 5-min interval at which the maximum number of steps is clocked is `r hiint`.

####Task 4: Impute missing values
```{r echo=TRUE}
#a. Calculate the total number of days/intervals where there are missing values.
sum(is.na(data[,1]))

#b. Strategy for populating the missing values = Use the mean for that 5-minute interval. 

#c. Create new dataset with missing data filled in
data_new = data
for(i in 1:nrow(data)){
        if(is.na(data[i,1])){
                data_new[i,1]=intervaldata[match(data[i,3],intervaldata[,1]),2]
        }
}

#d. Make a histogram of the total number of steps taken each day. Calculate the mean and median total number of steps taken per day.
daysteps_new <- sapply(grp, function(x) sum(data_new[x:(x+n-1),1]))
daydata_new <- matrix(nrow=length(grp),ncol=2)
colnames(daydata_new)<-c("total day's steps","date")
for(i in 1:length(grp)){
        daydata_new[i,1] <- daysteps_new[i]
        daydata_new[i,2] <- as.character(data_new[grp[i],2])
}
hist(as.numeric(daydata_new),xlab="",main="Histogram of the total number of steps taken per day")
meandaysteps_new = round(mean(daysteps_new,na.rm=TRUE),digits=3)
mediandaysteps_new=median(daysteps_new,na.rm=TRUE)
```

The new mean value is `r meandaysteps_new`, compared with the old value of `r meandaysteps`. The new median value is `r mediandaysteps_new`, compared with the old value of `r mediandaysteps`. It is clear that these are drastically different. Imputing missing data based on the average steps for that 5-min interval had the effect of drastically boosting the mean and median values - more so for the mean value.

```{r echo=TRUE}
#a. Create dataset with new factor variable indicating whether it's a weekday or weekend
data_new2=matrix(nrow=nrow(data_new),ncol=4)
data_new2[,1]=data_new[,1]
data_new2[,3]=data_new[,3]
data_new2[,2]=as.character(data_new[,2])
for(i in 1:nrow(data_new2)){
        if(weekdays(as.Date(data_new2[i,2]))=="Saturday"|weekdays(as.Date(data_new2[i,2]))=="Sunday"){
                data_new2[i,4]="Weekend"     
        } else {
                data_new2[i,4]="Weekday"
        }
}

#b. Create panel plot. The first plot is of the average steps taken in each 5-min interval, for weekdays. The second plot is same, but done for weekends.
intervaldata2 <- matrix(nrow=n,ncol=3)
colnames(intervaldata2)<-c("interval","average steps at this interval, weekdays","average steps at this interval, weekends")
j=1
k=1
grp_weekday=vector()
grp_weekend=vector()
for(i in 1:length(grp)){
        if(data_new2[grp[i],4]=="Weekday"){
                grp_weekday[j]=grp[i]
                j=j+1
        } else if(data_new2[grp[i],4]=="Weekend"){
                grp_weekend[k]=grp[i]
                k=k+1
        }       
}
for(i in 1:288){
        intervaldata2[i,1]=data[i,3]
        intervaldata2[i,2]=sum(data[grp_weekday[]+i-1,1],na.rm=TRUE)
        intervaldata2[i,3]=sum(data[grp_weekend[]+i-1,1],na.rm=TRUE)
}
par(mfrow=c(1,2),cex.main=0.9)
plot(intervaldata2[,1],intervaldata2[,2],type="l",xlab="Interval",ylab="Average steps",main="Average steps at each Interval, Weekdays")
plot(intervaldata2[,1],intervaldata2[,3],type="l",xlab="Interval",ylab="Average steps",main="Average steps at each Interval, Weekends")
```